DATA 605 Week 7 Assignment

### 2019-03-16 DATA605/MSilva_Assign7.Rmd
More work on Assignment 7

Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y .

@@

```{r echo=FALSE}
library(ggplot2)
n_failures_before_success <- 8
probability_of_success <- 1/10

@@

```{r}
expected_value <- 1 / probability_of_success

message(paste("Expected value:", expected_value))
```

@@

$\text{Standard Deviation} = \sqrt{\frac{1-P(success)}{P(success)^2}} = \sqrt{\frac{1-0.1}{0.1^2}} = \sqrt{\frac{0.9}{0.01}} = \sqrt{90} \approx 9.486833$

@@

standard_deviation <- sqrt(variance)

message(paste("Standard deviation:", standard_deviation))
```

@@

```{r}
p_exponential <- exp(1) ^ (-n_failures_before_success /expected_value)

@@

```{r}
expected_value <- 1 / probability_of_success

message(paste("Expected value:", expected_value))
```

@@

$\text{Standard Deviation} = \frac{1}{P(success)^2} = \frac{1}{0.1^2} = \frac{1}{0.01} = 100$

```{r}
standard_deviation <- 1 / (probability_of_success ^ 2)

message(paste("Standard deviation:", standard_deviation))
```

@@

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a binomial.**

@@

## Summary

|        | Probability of Failure After 8 Years | Expected Value | Standard Deviation |
| Part A | 0.4304672                            | 10             | 9.48683298050514   |
| Part B | 0.449329                             | 10             | 100                |
| Part C |
| Part D |

>>>>

Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y.

```{r}
set.seed(42)
n <- 100000
k <- 50
draws <- 10
Y <- c()
for (i in 1:n){
  Xn <- sample(1:k, draws, TRUE)
  Y <- c(Y, min(Xn))
}
```

```{r, echo=FALSE}
library(ggplot2)

ggplot(data.frame(table(Y)), aes(Y, Freq)) + 
  geom_bar(stat = "identity", fill = "steelblue", width = 1) +
  theme_minimal() +
  ylab("Count") + 
  ggtitle("Distribution of Y's")
```

It's a long tail distribution like a power law distribution.

@@

```{r echo=FALSE}
n_failures_before_success <- 8
probability_of_success <- 1/10

@@

```{r}
expected_value_geom <- 1 / probability_of_success

message(paste("Expected value:", expected_value_geom))
```

@@

$$\text{Standard Deviation} = \sqrt{\frac{1-P(success)}{P(success)^2}} = \sqrt{\frac{1-0.1}{0.1^2}} = \sqrt{\frac{0.9}{0.01}} = \sqrt{90} \approx 9.486833$$

@@

standard_deviation_geom <- sqrt(variance)

message(paste("Standard deviation:", standard_deviation_geom))
```

@@

```{r}
p_exponential <- exp(1) ^ (-n_failures_before_success /expected_value_geom)

@@

```{r}
expected_value_exp <- 1 / probability_of_success

message(paste("Expected value:", expected_value_exp))
```

@@

$$\text{Standard Deviation} = \sqrt{\frac{1}{P(success)^2}} = \sqrt{\frac{1}{0.1^2}} = \sqrt{\frac{1}{0.01}} = \sqrt{100} = 10$$

```{r}
standard_deviation_exp <- sqrt(1 / (probability_of_success ^ 2))

message(paste("Standard deviation:", standard_deviation_exp))
```

@@

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)**

The binomial probability formula is

$$P(x) = \frac{N!}{x!(N-x)!}\pi^x(1-\pi)^{N-x}$$

The probability of the maching failing after eight years is 1 minus the probability it does not fail in the first 8 years.  Otherwise put 0 successes(x) in the first 8 years (N).  The probability of failure is 1/10 therefore:

$$P(0) = \frac{8!}{0!(8-0)!}\frac{1}{10}^0(1-\frac{1}{10})^{8-0} = \frac{40320}{40320}\times 1 \times \frac{9}{10}^8 = 1 \times 1 \times 0.9^8 = 0.4304672$$

```{r}
p_binomial <- pbinom(0, 8, 0.1)
message(paste("Probability that the machine will fail after 8 years:", p_binomial))
```

The [expected value](http://onlinestatbook.com/2/probability/binomial.html) is the number of trials times the probability of success per trial.

```{r}
expected_value_binom <- n_failures_before_success * probability_of_success

message(paste("Expected value:", expected_value_binom))
```

The [standard deviation](http://onlinestatbook.com/2/probability/binomial.html) is found using:

$$\sigma = \sqrt{N\pi(1-\pi)}$$

```{r}
standard_deviation_binom <- sqrt(n_failures_before_success * probability_of_success*(1 - probability_of_success))

message(paste("Standard deviation:", standard_deviation_binom))
```

@@

The Poisson formula is:

$$P(x) = \frac{\lambda^xe^{-\lambda}}{x!}= \frac{10^8e^{-10}}{8!} =\frac{4539.993}{40320} = 0.112599$$

```{r}
x <- 8
mu <- 10
p_poisson <- ((mu^x) * (exp(1)^-mu)) / factorial(x)
message(paste("Probability that the machine will fail after 8 years:", p_poisson))
```

For the Poisson distribution the mean and standard deviation are the same.

```{r}
expected_value_poisson <- mu

message(paste("Expected value:", expected_value_poisson))
```

The variance is equal to mu so

```{r}
standard_deviation_poisson <- sqrt(mu)
```

### Summary

The following table summarizes all the information found above

```{r echo=FALSE, message=FALSE}
library(kableExtra)
library(dplyr)

data.frame(Model = c("Geometric", "Exponential", "Binomial", "Poisson"),
           Probability = c(p_geometric, p_exponential, p_binomial, p_poisson),
           ex = c(expected_value_geom, expected_value_exp, expected_value_binom, expected_value_poisson),
           sd = c(standard_deviation_geom, standard_deviation_exp, standard_deviation_binom, standard_deviation_poisson)) %>%
  rename("Expected Value" = ex,
         "Standard Deviation" = sd) %>%
  kable() %>%
  kable_styling()
```
